Environmental drivers of synchrony and their interactions

### 2022-11-08 Paper/Paper.Rnw
mainly added a bunch of comments which suggest how to restructure the write-up for simplicity, which is what I am going to do next

One of the classically theorized \citep{Moran1953} and now widely empirically validated mechanisms of 
synchrony, called the Moran effect, operates when spatially synchronized environmental signals, such
as temperature or precipitation, help drive population dynamics and thereby induce synchrony in
populations. Two recent papers have documented that Moran effects of distinct 
environmental drivers can, in fact, interact, either synergistically or antagonistically,
so that total population synchrony can be either greater or less than if the distinct Moran 
drivers operated independently \citep{Sheppard2019,CastoraniSheppardKelp}. Effects can be strong.
\cite{Sheppard2019} showed that long-timescale ($>4$y period) synchrony in a chlorophyll density index 
in the seas around the United Kingdom is substantially augmented by interactions between environmental 
drivers of synchrony and drivers linked with consumption by a copepod consumer.
\cite{CastoraniSheppardKelp} showed that nutrient dynamics and wave disturbance, two Moran drivers
of synchrony in giant kelp populations on the California (CA) coast, interact either synergistically 
or destructively, depending on which portion of the CA coast is examined and on 
timescale of analysis. 
Both of these studies argued that interactions between Moran effects may be a general feature of many 
systems, because of the large number of interrelated factors driving population fluctuations in most 
spatially extended systems. The main purpose of this study is to establish a general theory of the mechanism
of interacting Moran effects and to use examples and applications of the theory to provide 
ecological intuition for how the mechanism of interacting Moran effects works. Our goals are
distinct from those of \cite{Kendall2000}, who examined interactions between Moran and dispersal
causes of synchrony.

@@

Having demonstrated that interactions between Moran effects can augment or reduce population synchrony, we now 
begin to provide some intuition on how to understand the underlying mechanism. 
A key insight is the use of a timescale-specific approach,
i.e., environmental signals are decomposed, using Fourier analysis, into components that occur on different timescales,
and the nature of interacting Moran effects can be understood on a timescale-by-timescale basis, with effects summed 
across timescales afterwards if desired. Fig. \ref{fig:PedagogAlt}a,b shows the period-20 Fourier components of two hypothetical,
spatially synchronous environmental variables measured in sampling locations 
$i=1,2,3$, and one way these can influence populations. The components are
lagged, relative to each other, in the timing of their peaks ($l_n$ on the figure). They are also lagged in their 
effects on populations, by the amounts $l_{e1}$ and $l_{e2}$, respectively, i.e., peaks in the environmental signals
manifest as maximum positive influence on populations after delays of $l_{e1}$ and $l_{e2}$, respectively. In this example,
we assume for simplicity that larger values of both environmental variables are beneficial to populations, though the general
theory described below does not require that assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_n+l_{e1}=l_{e2}$, then the 
periodic maximal positive influences of the two environmental variables coincide with each other as well as being 
spatial synchronous. This alignment of produces additional synchrony in populations, 
beyond what would manifest if environmental fluctuations were unrelated, because positive influences of variable 1 in location $i$ will
tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:PedagogAlt}c, d
shows the opposite scenario, for which $l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period. Thus
the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
influence of environmental variable 2, reducing synchrony relative to the case of unrelated environmental variables. 
Intermediate scenarios are possible, as will be revealed by the general theory.
The theory also formalizes the timescale decomposition of synchrony measures, and timescale-specific lags.

Our specific goals for this study are as follows. First, (G1) we will elaborate a general mathematical 
theory of interacting Moran effects which allows a detailed understanding of how the mechanism works. Second, 
(G2) we will consider theoretical case studies, which emerge as important special cases of our general theory, and 
which illuminate the intuition behind how Moran effects may interact in real systems. Third, (G3) we will 
apply our theory to populations of giant kelp off the CA coast. Whereas \cite{CastoraniSheppardKelp}
already carried out a detailed analysis of interacting Moran effects in CA kelp populations and their importance 
for kelp ecology, we instead use a simplified subset of the available kelp data, and our analysis is instead intended 
to illuminate the inner workings of the general mechanism we describe. In the Discussion, we 
extend the arguments of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} that interactions between Moran 
effects are likely to be common, and hence that our general theory seems likely to pertain to numerous 
ecosystems. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
illuminate this newly observed but potentially quite important mechanism of spatiotemporal population dynamics.

@@

Comparing the terms on the right
of (\ref{rhoww_T1to3}) gives the relative contributions of direct Moran effects and interactions 
between Moran effects. 
The magnitudes of the quantities $f_{\mcP^{(1)}}$ and $f_{\mcP^{(2)}}$ quantify the strength of 
influence of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ on populations at the timescale $\sigma=1/f$. 
The phases of these quantities encode the lags in the influences of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$, 
represented on Fig. \ref{fig:PedagogAlt} by $l_{e1}$ and $l_{e2}$,
relative to the timescale $\sigma$. The lag 
$l_n$ on Fig. \ref{fig:PedagogAlt} manifests in the theory 
through the phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$. 
The expression $l_{e1}-l_{e2}+l_{n}$, which quantifies the degree of mismatch among the lags
(see Introduction and Fig. \ref{fig:PedagogAlt}) corresponds to the phase of the expression 
$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, because the phase of
this product equals the phase of $f_{\mcP^{(1)}}$ (which corresponds to $l_{e1}$), minus 
the phase of $f_{\mcP^{(2)}}$
(which corresponds to $l_{e2}$), plus the phase
of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ (which corresponds to $l_n$). 
The real part of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
is positive and larger whenever the phase of this quantity is close to zero, i.e., when    
$l_{e1}-l_{e2}+l_n$ is close to zero (Fig. \ref{fig:PedagogAlt}a, b), and is increasingly negative as 
the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ gets
close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to half the period of oscillation
(Fig. \ref{fig:PedagogAlt}c, d). The factor $1/|f_{\mcB}|^2$ is how the intrinsic nature of population
dynamics modulates Moran influences.
The derivation of the theory is in SI section \ref{SM-sect:derivation}.

@@

@@

To estimate the components $\rho_{ww}$, $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and 
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), where $w$ was linearly detrended kelp time series in one of our regions, 
$\epsilon^{(1)}$ was detrended nitrate time
series in the region, and $\epsilon^{(2)}$ was detrended wave disturbance time series in the region, 
we applied the definitions of these quantities (see General Theory), which required estimation from data of
the spectra and cross spectra $S_{w_i w_j}$ and $S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}$ for $i,j=1,\ldots,N$
and $a,b=1,2$, where $N$ is the number of locations in the region being considered. Spectral quantities were
computed using the consistent estimator of section 7.4 of \cite{Brillinger2001}. 
The estimator is a smoothed periodogram, with the width of the smoothing kernel selected to 
increase with the square root of time series length. Theory was interpreted in relation to kelp ecology and
theoretical case studies by plotting the components of (\ref{rhoww_T1to3}) for each of our regions.

@@

%Para basically confirming earlier work, as a starter
To begin fulfilling goal G2 of the Introduction, our theoretical case studies 
demonstrate that interaction effects between Moran drivers: 1) can be 
comparable in strength to direct Moran effects; 2) can be either synergistic or destructive; and 3) 
can depend strongly on timescale. 
First, for all of our case studies, the magnitude of 
interaction effects was comparable to that of direct Moran effects (Fig. \ref{fig:TheoryCaseA}a,e,i,m, 
compare the dashed and solid lines). Thus interactions can contribute substantially to overall 
synchrony. Second, in contrast to direct Moran effects, which are positive, 
interactions can be negative or positive. CaseA and CaseB
showed negative interaction effects on short timescales (Fig. \ref{fig:TheoryCaseA}a,e); 
CaseC1 showed negative interactions on
long timescales (Fig. \ref{fig:TheoryCaseA}i); and CaseC2 showed negative interactions on all timescales 
(Fig. \ref{fig:TheoryCaseA}m).
Thus interaction effects can either augment or reduce synchrony. Finally, interaction effects depended strongly on timescale 
for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and even differed in sign across timescales for CaseA, CaseB and CaseC1. 
This result complements earlier studies that showed direct Moran effects can 
depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The results of this paragraph
also follow from \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not 
straightforwardly. And those papers are case studies, so our results provide novel theoretical generality.

%Case study A, new try
CaseA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
Moran effects on different timescales, and how our theory captures that contrast.
For CaseA, $\epsilon^{(1)}$ effects on populations were lagged by $1$q but 
$\epsilon^{(2)}$ effects were unlagged (Methods). So, in the language of Fig. \ref{fig:PedagogAlt},
$l_{e1}=1$q and $l_{e2}=0$q. 
The between-noise lag, $l_n$, of Fig. \ref{fig:PedagogAlt} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details, here).
Thus, what determines the sign of interactions between Moran effects for this example is how 
$l_{e1}-l_{e2}+l_n=1$q compares to the timescale/period, $\sigma$, being considered. 
On the shortest timescales ($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}a, left side of panel), $1$q
is half the period, so interactions effects are negative. On long timescales (e.g., $\sigma>8$y), $1$q is a negligible portion of the
period, so $l_{e1}-l_{e2}+l_n=1$q is close to zero, relative to that period, and interaction effects
are positive: compared to such long timescales, $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects happen close to simultaneously, so
the two noise variables reinforce each other. Theory captures these complexities because
the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, which equals the phase of
$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ because $l_n=0$ and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number
(SI section \ref{SM-sect:examples_cross_sync}). But $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=
p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
which has phase $-\pi$ at short timescales
and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
interaction term in (\ref{rhoww_T1to3}), as described in General Theory. Comparing
Fig. \ref{fig:TheoryCaseA}a and d shows the additional influence of autoregressive population effects, which
simply multiply all Moran influences by the same timescale-dependent non-negative 
quantity, not altering their relative importance or sign [see (\ref{rhoww_T1to3})].

%Case study B, new version
CaseB reveals how Moran effects can interact when lags are longer than 1 model time step. 
For CaseB, $\epsilon^{(1)}$ effects on populations were lagged by $n$q but 
$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$, so $l_{e1}=3$q and $l_{e2}=0$q. 
As for CaseA, because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
a white noise process for CaseB, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ was a real number and $l_n=0$.
Interaction effects were again negative on short timescales 
($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}e, far left side of panel)
because $l_{e1}-l_{e2}+l_n=3$q was $1.5\sigma$ on that timescale, and so effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ were in a half-phase relationship and
counteracted each other.
Similar to CaseA, $l_{e1}-l_{e2}+l_n=3$q was again negligible compared to long timescales, so interaction effects
were positive on long timescales (Fig. \ref{fig:TheoryCaseA}e, right side of panel).
The quantity $l_{e1}-l_{e2}+l_n=3$q exactly equaled the timescale for $\sigma=3 \text{q} = 0.75 \text{y}$ and 
equaled half the timescale for $\sigma=6 \text{q} = 1.5 \text{y}$, hence interactions effects
were positive for $\sigma=0.75 \text{y}$ and negative for $\sigma=1.5$y 
(Fig. \ref{fig:TheoryCaseA}e). Theory captures these complexities because 
the size of the $3$q sum of lags for this scenario, relative to $\sigma$,
is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$; the phase of that
quantity, represented by colors on Fig. \ref{fig:TheoryCaseA}f,g, fluctuates analogously to the
interaction effects on Fig. \ref{fig:TheoryCaseA}a.

>>>>

One of the classically theorized \citep{Moran1953} and now widely empirically validated mechanisms of 
synchrony, called the Moran effect, operates when spatially synchronized environmental signals, such
as temperature or precipitation, help drive population dynamics and thereby induce synchrony in
populations. Two recent papers have documented that Moran effects of distinct 
environmental drivers can, in fact, interact, either synergistically or antagonistically,
so that total population synchrony can be either greater or less than if the distinct Moran 
drivers operated independently \citep{Sheppard2019,CastoraniSheppardKelp}. Effects can be strong.
\cite{Sheppard2019} showed that long-timescale ($>4$y period) synchrony in a chlorophyll density index 
in the seas around the United Kingdom is substantially augmented by interactions between environmental 
drivers of synchrony and drivers linked with consumption by a copepod consumer.
\cite{CastoraniSheppardKelp} showed that nutrient dynamics and wave disturbance, two Moran drivers
of synchrony in giant kelp populations on the California (CA) coast, interact either synergistically 
or destructively, depending on which portion of the CA coast is examined and on 
the timescale of analysis. 
Both of these studies argued that interactions between Moran effects may be a general feature of many 
systems, because of the large number of interrelated factors driving population fluctuations in most 
spatially extended systems. The main purpose of this study is to establish a general theory of the mechanism
of interacting Moran effects and to use examples and applications of the theory to provide 
ecological intuition for how the mechanism of interacting Moran effects works. Our goals are
distinct from those of \cite{Kendall2000}, who examined interactions between Moran and dispersal
causes of synchrony.

@@

%***DAN: The following para is a lot to ask the reader to absorb in the Intro, way too much for 
%an ecological audience. Move it Theory. Can I replace it with something more intuitive, like the
%swing example? Or is that a bad combination of too involved and too informal/non-ecological? 
%If I were to use the swing example I'd have to do it in limited space, to prevent it from 
%spiraling out of control as an overly extended metaphor.
%***DAN: You also might want to add a para to the Intro on why interactions between Moran effects
%should be important, to take up some space upon moving the below para to Theory (and it would also 
%help motivate).
Having demonstrated that interactions between Moran effects can augment or reduce population synchrony, we now 
begin to provide some intuition on how to understand the underlying mechanism. 
A key insight is the use of a timescale-specific approach,
i.e., environmental signals are decomposed, using Fourier analysis, into components that occur on different timescales,
and the nature of interacting Moran effects can be understood on a timescale-by-timescale basis, with effects summed 
across timescales afterwards if desired. Fig. \ref{fig:PedagogAlt}a,b shows the period-20 Fourier components of two hypothetical,
spatially synchronous environmental variables measured in sampling locations 
$i=1,2,3$, and one way these can influence populations. The components are
lagged, relative to each other, in the timing of their peaks ($l_n$ on the figure). They are also lagged in their 
effects on populations, by the amounts $l_{e1}$ and $l_{e2}$, respectively, i.e., peaks in the environmental signals
manifest as maximum positive influence on populations after delays of $l_{e1}$ and $l_{e2}$, respectively. In this example,
we assume for simplicity that larger values of both environmental variables are beneficial to populations, though the general
theory described below does not require that assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_n+l_{e1}=l_{e2}$, then the 
periodic maximal positive influences of the two environmental variables coincide with each other as well as being 
spatial synchronous. This alignment of influences produces additional synchrony in populations, 
beyond what would manifest if environmental fluctuations were unrelated, because positive influences of variable 1 in location $i$ will
tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:PedagogAlt}c, d
shows the opposite scenario, for which $l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period. Thus
the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
influence of environmental variable 2, reducing synchrony relative to the case of unrelated environmental variables. 
Intermediate scenarios are possible, as will be revealed by the general theory.
The theory also formalizes the timescale decomposition of synchrony measures, and timescale-specific lags.

Our specific goals for this study are as follows. First, (G1) we will elaborate a general 
theory of interacting Moran effects which allows a detailed understanding of how the mechanism works. Second, 
(G2) we will consider theoretical case studies, which emerge as important special cases of our general theory, and 
which illuminate the intuition behind how Moran effects may interact in real systems. Third, (G3) we will 
apply our theory to populations of giant kelp off the CA coast. Whereas \cite{CastoraniSheppardKelp}
already carried out a detailed analysis of interacting Moran effects in CA kelp populations and their importance 
for kelp ecology, we instead use a simplified subset of the available kelp data. Our analysis is intended 
to illuminate the inner workings of the general mechanism we describe rather than being a complete exploration
of kelp dynamics. In the Discussion, we 
extend the arguments of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} that interactions between Moran 
effects are likely to be common, and hence that our theory seems likely to pertain to numerous 
ecosystems. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
illuminate this newly observed but potentially quite important mechanism of spatiotemporal population dynamics.

@@

Comparing the terms on the right
of (\ref{rhoww_T1to3}) gives the relative contributions of direct Moran effects and interactions 
between Moran effects. 
The magnitudes of the quantities $f_{\mcP^{(1)}}$ and $f_{\mcP^{(2)}}$ quantify the strength of 
influence of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ on populations at the timescale $\sigma=1/f$. 
The phases of these quantities encode the lags in the influences of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$, 
represented on Fig. \ref{fig:PedagogAlt} by $l_{e1}$ and $l_{e2}$,
relative to the timescale $\sigma$. The lag 
$l_n$ on Fig. \ref{fig:PedagogAlt} manifests in the theory 
through the phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$. 
The expression $l_{e1}-l_{e2}+l_{n}$, which quantifies the degree of mismatch among the lags
(see Introduction and Fig. \ref{fig:PedagogAlt}) corresponds to the phase of the expression 
$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, because the phase of
this product equals the phase of $f_{\mcP^{(1)}}$ (which corresponds to $l_{e1}$), minus 
the phase of $f_{\mcP^{(2)}}$
(which corresponds to $l_{e2}$), plus the phase
of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ (which corresponds to $l_n$). 
The real part of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
is positive and larger whenever the phase of this quantity is close to zero, i.e., when    
$l_{e1}-l_{e2}+l_n$ is close to zero (Fig. \ref{fig:PedagogAlt}a, b), and is increasingly negative as 
the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ gets
close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to half the period of oscillation
(Fig. \ref{fig:PedagogAlt}c, d). The factor $1/|f_{\mcB}|^2$ is how the intrinsic nature of population
dynamics modulates Moran influences.
The derivation of the theory is in SI section \ref{SM-sect:derivation}.
%***DAN: Some of the content of the above para is quite technical, and is more about interpreting the
%inner working of the theory as opposed to just comparing its predictions to the intution from
%Figs 2 and 5. So maybe some of this should be move to a sup mat where I would really
%delve into how the inner workings of the theory correspond to all these lags and alignments
%of lags and so on.

@@

%***DAN: Consider renaming our three scenarios as follows: A->A1, B->A2, C1->B1, C2->B2. After all,
%A and B are quite related, so it makes sense to consider the as scenarios within a case study.
%Then it would be more symmetric, and the four rows of panels of Fig. 4 would each correspond to
%a scenario within a case study.

@@

To estimate the components $\rho_{ww}$, $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and 
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), where $w$ was linearly detrended kelp time series in one of our regions, 
%***DAN: Probably should shift this definition of what w stands for to the previous paragraph, since I was 
%already using w in this context, there.
$\epsilon^{(1)}$ was detrended nitrate time
series in the region, and $\epsilon^{(2)}$ was detrended wave disturbance time series in the region, 
%***DAN: likewise these ought to have been defined in the previous paragraph, I think
we applied the definitions of these quantities (see General Theory), which required estimation from data of
the spectra and cross spectra $S_{w_i w_j}$ and $S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}$ for $i,j=1,\ldots,N$
and $a,b=1,2$, where $N$ is the number of locations in the region being considered. Spectral quantities were
computed using the consistent estimator of section 7.4 of \cite{Brillinger2001}. 
The estimator is a smoothed periodogram, with the width of the smoothing kernel selected to 
increase with the square root of time series length. Theory was interpreted in relation to kelp ecology and
theoretical case studies by plotting the components of (\ref{rhoww_T1to3}) for each of our regions.

@@

%Para basically confirming earlier work, as a starter
To begin fulfilling goal G2 of the Introduction, our theoretical case studies 
demonstrate that interaction effects between Moran drivers: 1) can be 
comparable in strength to direct Moran effects; 2) can be either synergistic or destructive; and 3) 
can depend strongly on timescale. 
First, for all of our case studies, the magnitude of 
interaction effects was comparable to that of direct Moran effects (Fig. \ref{fig:TheoryCaseA}a,e,i,m, 
compare the dashed and solid lines). Thus interactions can contribute substantially to overall 
synchrony. Second, in contrast to direct Moran effects, which are positive, 
interactions can be negative or positive. CaseA and CaseB
showed negative interaction effects on short timescales (Fig. \ref{fig:TheoryCaseA}a,e); 
CaseC1 showed negative interactions on
long timescales (Fig. \ref{fig:TheoryCaseA}i); and CaseC2 showed negative interactions on all timescales 
(Fig. \ref{fig:TheoryCaseA}m).
Thus interaction effects can either augment or reduce synchrony. Finally, interaction effects depended strongly on timescale 
for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and even differed in sign across timescales for CaseA, CaseB and CaseC1. 
This result complements earlier studies that showed direct Moran effects can 
depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The results of this paragraph
also follow from \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not 
straightforwardly. And those papers are case studies, whereas our results provide general theory.

%Case study A, new try
CaseA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
Moran effects on different timescales, and how our theory captures that contrast.
For CaseA, $\epsilon^{(1)}$ effects on populations were lagged by $1$q but 
$\epsilon^{(2)}$ effects were unlagged (Methods). So, in the language of Fig. \ref{fig:PedagogAlt},
$l_{e1}=1$q and $l_{e2}=0$q. 
The between-noise lag, $l_n$, of Fig. \ref{fig:PedagogAlt} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details, here).
Thus, what determines the sign of interactions between Moran effects for this example is how 
$l_{e1}-l_{e2}+l_n=1$q compares to the timescale/period, $\sigma$, being considered. 
On the shortest timescales ($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}a, left side of panel), $1$q
is half the period, so interaction effects are negative. On long timescales (e.g., $\sigma>8$y), $1$q is a negligible portion of the
period, so $l_{e1}-l_{e2}+l_n=1$q is close to zero, relative to that period, and interaction effects
are positive: compared to such long timescales, $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects happen close to simultaneously, so
the two noise variables reinforce each other. 
%***DAN: Upon re-reading aftre taking several months break from this project, I find myself skimming
%the following sentences which are supposed to be showing how the theory captures what is going on.
%It makes me wonder if explaining how theory corresponds to intuition is maybe beyond the audience,
%and that part should be consistent relegated to sup mat, and instead in the main text we just 
%present the results of theory, and explain how they correspond to the intuition developed in Fig. 2.
%I think that would also allow us to put Fig. 4 panels b,c,f,g,j,k,n,o into sup mat, massively
%simplifying the presentation. Keep the last sentence of this para here, though, along with panels 
%d,h,l,p.
Theory captures these complexities because
the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, which equals the phase of
$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ because $l_n=0$ and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number
(SI section \ref{SM-sect:examples_cross_sync}). But $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=
p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
which has phase $-\pi$ at short timescales
and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
interaction term in (\ref{rhoww_T1to3}), as described in General Theory. Comparing
Fig. \ref{fig:TheoryCaseA}a and d shows the additional influence of autoregressive population effects, which
simply multiply all Moran influences by the same timescale-dependent non-negative 
quantity, not altering their relative importance or sign [see (\ref{rhoww_T1to3})].

%Case study B, new version
%***DAN: Note that case B was introduced in Methods as having a lag of length n in the effects
%of epsilon 1, but we immediate specify n=3 in Results. Do we want to instead intro it as n=3
%from the beginning?
CaseB reveals how Moran effects can interact when lags are longer than 1 model time step. 
For CaseB, $\epsilon^{(1)}$ effects on populations were lagged by $n$q but 
$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$, so $l_{e1}=3$q and $l_{e2}=0$q. 
As for CaseA, because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
a white noise process for CaseB, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ was a real number and $l_n=0$.
Interaction effects were again negative on short timescales 
($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}e, far left side of panel)
because $l_{e1}-l_{e2}+l_n=3$q was $1.5\sigma$ on that timescale, and so effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ were in a half-phase relationship and
counteracted each other.
Similar to CaseA, $l_{e1}-l_{e2}+l_n=3$q was again negligible compared to long timescales, so interaction effects
were positive on long timescales (Fig. \ref{fig:TheoryCaseA}e, right side of panel).
The quantity $l_{e1}-l_{e2}+l_n=3$q exactly equaled the timescale for $\sigma=3 \text{q} = 0.75 \text{y}$ and 
equaled half the timescale for $\sigma=6 \text{q} = 1.5 \text{y}$, hence interactions effects
were positive for $\sigma=0.75 \text{y}$ and negative for $\sigma=1.5$y 
(Fig. \ref{fig:TheoryCaseA}e). 
%***DAN: We could again put this "theory captures these complexities" business in sup mat, where
%we could anyway explain in greater detail. Also put the middle panels there.
Theory captures these complexities because 
the size of the $3$q sum of lags for this scenario, relative to $\sigma$,
is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$; the phase of that
quantity, represented by colors on Fig. \ref{fig:TheoryCaseA}f,g, fluctuates analogously to the
interaction effects on Fig. \ref{fig:TheoryCaseA}a.

%Case study C, new
CaseA and CaseB assumed white noise processes, and therefore 
real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ and $l_n=0$.
But CaseC illustrates what can happen when noise processes have periodic components and lagged interdependences, e.g., 
both processes oscillate with annual periodicity and distinct phenology, an important scenario because seasonality is common.
For CaseC, $\epsilon^{(1)}$ effects on populations were lagged by $n_1$q and 
$\epsilon^{(2)}$ effects were lagged by $n_2$q (Methods). 
%***DAN: give the exact lag values used in methods. Also the coefficient values, probably. 
We used $n_1=1$, $n_2=0$, $p_1^{(1)}=3$ and $p_0^{(2)}=-1.5$, so 
$\epsilon^{(1)}$ effects were lagged and positive, and $\epsilon^{(2)}$ effects 
were immediate and negative. The language and logic of 
Fig. \ref{fig:PedagogAlt} does not directly apply, because we there assumed, in order to introduce concepts simply,
that environmental effects were positive, and we now take $p_0^{(2)}=-1.5$. But similar logic does apply 
(Fig. \ref{fig:PedagogAlt_SM}), with
both the new and previous paradigms covered by our general theory.
%***DAN: It is asking way too much of the reader to spend the labor to go through Fig. 5 and update their
%understanding to become more general than the original Fig. 2 just because of this one sentence. I am
%asking them to spend maybe 15-20 minutes puzzling over Fig. 5 based on this one sentence of main text.
%People won't do that, instead they'll just accept a more surface-level understanding and blunder onward.
%Suppose I moved the text referring to Fig. 2 from the Intro to the Theory, and then also cited what is
%currently Fig. 5 there (it would become Fig. 3). So the reader would be prepared for both positive and
%negative effects, and when they got here they would be ready to absorb this example. I think this is a 
%smart idea.
For CaseC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
$1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}). Thus lags compounded: the 
annual positive population effects of $\epsilon^{(1)}$ 
were exactly $2$q offset from the annual negative effects of $\epsilon^{(2)}$, 
producing reinforcing interactions 
between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i).
Contrastingly, for CaseC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
$1$q. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
coincided with the negative effects of $\epsilon^{(2)}$ every year, producing 
negative interactions between Moran influences (Fig. \ref{fig:TheoryCaseA}m).
%***DAN: More "theory captures these complexities" stuff that should be sent to sup mat.
%No one will absorb this kind of stuff who isn't incredibly dedicated, and therefore
%will be willing to read sup mat. Make the main text more about comparing the intution 
%from fig 2 and 5 with the predictions of theory, while black-boxing how the theory
%actually works.
Our theory captures these complexities
because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
quantifies the degree of alignment between the effects of $\epsilon^{(1)}$ and 
$\epsilon^{(2)}$ compared to a given timescale, while 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the degree of alignment between 
oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
therefore captures whether delays in effects and lags between 
variables align or misalign, with alignment
reflected in the real part of this quantity, which is half the interaction term in our theory (\ref{rhoww_T1to3}).
On long timescales, interactions were the
same in the two scenarios, and were negative, because, on those timescales, sub-annual lags 
make negligible difference, and the effects of $\epsilon^{(1)}$ on populations were 
positive and those of $\epsilon^{(2)}$ were negative.
